---
title: "Exploring SportsCode XMLs"
author: "James Rogol"
date: "`r Sys.time()`"
output: html_document
---

## Loading data
We'll use the `tidyverse` and `xml2` packages for parsing the XML. A dummy XML was exported from SportsCode using `File > Export > Sportstec XML...`

```{r ingest, message=FALSE, warning=FALSE}
library(tidyverse)
library(xml2)

xml_file <- "test_data/exported.xml"

xml <- xml2::read_xml(xml_file)
```

## Structure
### <ALL_INSTANCES>
There are 2 items in the resulting XML list `<ALL_INSTANCES>` and `<ROWS>`. The first contains the metadata for each instance in the timeline. For simplicity's sake, we'll examine a single instance, using the XPATH `.//instance` (all nodes below the current one named "instance").
```{r all_instances}
xml_structure(xml_find_first(xml,".//instance"))
```
This consists of:

  * `instance`: The parent node for the following nodes.
  * `ID`: a unique identifier for each instances, assigned chronologically
  * `start`: start time of the instance, in seconds. `0` is the start of the timeline.
  * `end`: end time of the instance, in seconds.
  * `code`: The name of the row in SportsCode
  * `label`: XML parent node for individuals label. Multiple `label` nodes can exist for a given instance.
    * `group`: name of a label group, if it exists
    * `text`: the label's name.

All nodes contain text.

###<ROWS>
As `<ROWS>` is fairly short, we can look at the entire structure here.

```{r rows}
xml_structure(xml_find_first(xml,".//ROWS"))
```

  * `<row>`: Parent node for each row, similar to `<instance>`
  * `<code>`: The name of a row
  * `<R>`, `<B>`, `<G>`: contain RGB values for the color of the row, on a scale of 0-100.

## Digging into `<ALL_INSTANCES>`
We're not so much concerned with the data in `<ROWS>`, as that's more pertinent when re-importing data into SportsCode. As `<code>` nodes exist in `<ROWS>`, we'll extract `<ALL_INSTANCES>` to make our lives easier.[^1]
```{r}
instances <- xml %>% xml_find_first( "./ALL_INSTANCES")
```

### Basic Information
Extracting the basic information should be pretty straightforward. This includes everything outside of the `<label>` nodes, as there can be multiple occurrences, including some within the same named `group`.

As all of the nodes are text fields, `xml2::xml_text()` extracts the values, returning a vector. We can then cast the numeric values to the appropriate type.
```{r}
instances %>% xml_find_all(".//ID") %>% xml_text() %>% as.integer()
```
```{r}
instances %>% xml_find_all(".//code") %>% xml_text()
```
```{r}
instances %>% xml_find_all(".//start") %>% xml_text() %>% as.double()
```
```{r}
instances %>% xml_find_all(".//end") %>% xml_text() %>% as.double()
```

Hadley Wickham states that any time a process is performed more than thrice, create a function. We'll do just that, borrowing from Bob Rudis' [snippet](https://rpubs.com/hrbrmstr/xml2power).

```{r xtrct}
xtrct <- function(doc, name) {
  xml_find_all(doc, paste0(".//",name)) %>%
    xml_text() %>%
    trimws() %>% 
    list() %>% 
    set_names(name)
  }
```
`xtrct` takes two arguments: `doc`, the the XML to parse; and `name`, the name of the target node (as a string). The function then automatically creates the XPATH, concatenating `.//` and `name`. As above, all the nodes are found, and their text values parsed, trimming any white space. This vector is then turned into a named list, using `name`.

```{r}
xtrct(instances,"code")
```

We can then iterate over a list of the "basic" variables, iterating over it, and creating a nice little tibble.[^2] We can then perform type casting using `dplyr::mutate_at`.
```{r}
basic <- c("ID","code","start","end")

basic_df <- basic %>% map_dfc(~xtrct(instances,.x)) %>%
  mutate_at(c("start","end"), as.double) %>% 
  mutate_at("ID",as.integer)
  
basic_df
```

We can also introduce a `duration` column by subtracting `start` from `end`:
```{r}
basic_df %>% mutate(duration = end-start)
```

### Labels and Groups

To Do!

## Overlap

SportsCode features the `overlap(rowA,rowB)` function, which returns all instances in either `rowA` or `rowB` which overlap at any point in time. This query can be refined to show only instances from `rowA`:
```{r eval = FALSE}
overlap(rowA,rowB) and rowA
```

It would be nice to introduce this functionality to SportsCode. As of Feb. 2017, there is a known issue with `dplyr` and `S4` objects, whereby the incorrect results are returned when using certain verb-functions.[^3]



## Footnotes
[^1]: Using the XPATH selector `.//code` looks for all child nodes, irrespective of the current branch. As such, it would locate nodes in both `<ROWS>` and `<ALL_INSTANCES>`. Furthermore, `./code` will only look in the current branch, necessitating us to navigate to `<ALL_INSTANCES>` anyway.
[^2]: `xtrct` generates lists, which will become the columns. `map_dfc` creates the tibble column-wise, as opposed to row-wise.
[^3]: See the corresponding GitHub issue [here](https://github.com/tidyverse/dplyr/issues/2432)
